Conference Paper Title\\

### 2023-05-31 LARS_paper/paper.tex
Update robot development subsections

A small quadruped was designed and developed for research (Fig. X).
During the prototyping phase, the hardware architecture and physics structure was designed. Following that, the structure of the robot was printed and integrated with the hardware components. The robot's actuators are servo motors of the model dynamixel MX-28, and a Raspberry Pi 4 was chosen as processing unit. Furthermore, it contains an inertial sensor, model MPU6050. All the software was developed with the Robot Operating System 2 Humble (ROS 2) \cite{ROS2Humble}.

@@

The robot's structure is mammal-type, which is characterized by the presence of two joints in the sagittal plane, similar to animals such as dogs and horses.  The robot’s legs follow a configuration called full-elbow, in which the four legs are oriented backward, as shown in Fig. X. Furthermore, it features 3 degrees of freedom (DoF) per leg, providing the legs with substantial freedom of movement. The robot was designed to favor the mass balance between the body and the robot's legs, that is, the majority of its mass is in located in the body or as close to it as possible. Lighter legs can move quickly without significantly changing the center of gravity of the robot, which increases the stability and requires less control complexity.  On the other hand, the legs should be strong enough to support the robot's weight and the impact with the ground.

The internal components, which include sensors, the processing unit, and the communication interface with the actuators were arranged symmetrically in the robot's body. The actuators (components that contribute with most of the system's mass) were positioned as close as possible to the body. The motor that acts on the tibia joint was placed in the upper part of the femur, in order to reduce the moment of inertia of the leg. This choice required the addition of a transmission system between the engine and the tibia link, consisting of a rigid metal rod with a ball joint at each edge.

Quadruped robots move according to a gait pattern. The gait implemented for the robot follows the sequence of phases demonstrated in Fig. \ref{fig:trot_pattern}, in which the white rectangles represent the  swing phase and the gray rectangles represent the stance phase. During the stance phase, the legs are on the ground and push the robot in the desired direction.  In the swing phase, the legs are raised and moved to the next support point. This gait pattern is called a trot, and is based on the concept of periodic and symmetrical gaits. The periodic gaits are characterized by the continuous repetition of the movements in the same instants. The symmetry is a characteristic of gaits that move a pair of legs together, leaving and returning to the ground in a synchronized way. Although the structure of the robot makes it possible to perform other types of gaits, in this work, only the trot was considered. In order to reduce the control complexity, a discontinuous gait was adopted, that is, the robot's body moves only when all the legs are on the ground, resulting in an intermittent movement. In Fig. \ref{fig:trot_pattern}, it is noticeable that the same pair of diagonal legs always move at the same instant. Between two consecutive swing phases, there is a moment when all legs are in stance, which is when the robot body is displaced in the desired direction of locomotion.

@@

The control subsystem of the robot is composed of two main components: the body's stabilization controller and 
joints' controllers. The body's stabilization controller is responsible for controlling the pitch and roll angles of the robot, maintaining it at 0 degrees. Meanwhile, the joints' controllers, each one a PID, are responsible for controlling each joint of the robot independently.

The gait planner is responsible for controlling each leg and the robot's body. It defines where each foot will step and where the body should move to based on the desired velocity command. The gait planner uses a trajectory planner to calculate the trajectory each foot and body should follow. This trajectory is calculated based on the velocity command. The points of this trajectory are converted into the angles of the joints through the kinematic model. The kinematic model of the robot is responsible for mapping the three-dimensional position of the foot to the joints' angles of the respective leg.

The gait planner sends the points at which each foot must be at a frequency of $\SI{50}{\hertz}$. This frequency was defined as being the highest that the robot processing system was able to support and by respecting the Nyquist theorem, which indicates that in this case, the frequency should be greater than $\SI{4}{\hertz}$ --- since the step period is $0.5 s$.

% TODO ref nyquist

\subsection{Kinematic model}
\label{sec:detail_inv_kinematics}

@@

\subsection{Stabilization control}

Two identical PID controllers were employed to control the \textit{roll} and \textit{pitch} rotation of the robot's body. They act independently, stabilizing the rotation in each axis simultaneously. Both controllers are identical and were implemented following the block diagram in Fig. \ref{fig:pid}.

@@

The IMU sensor is responsible for measuring the rotation of the robot's body, providing feedback on the system's outputs. The limiter was added to avoid body rotations that exceed the limits of the joints. The stabilization controller's outputs are sent to the gait planner.

\subsection{Trajectory planner}

The trajectory planner is responsible for calculating the trajectory that each foot must perform. For both stance and swing phases, a cycloidal curve is adopted. As shown in \cite{Shi2021}, a cycloidal curve can be defined in the three-dimensional space by (\ref{eq:traj_x}) to (\ref{eq:traj_k})
between the points $(x_o, y_o, z_o)$ and $(x_f, y_f, z_f)$ as a function of time $t$, step height $H$, and period $T$.
\begin{equation}
  x = (x_f - x_o) \frac{K - \sin{(K)}}{2 \pi} + x_o
  \label{eq:traj_x}
\end{equation}
\begin{equation}
  y = (y_f - y_o) \frac{K - \sin{(K)}}{2 \pi} + y_o
  \label{eq:traj_y}
\end{equation}
\begin{equation}
  z = H \frac{1 - \cos{(K)}}{2} + z_o
  \label{eq:traj_z}
\end{equation}
\begin{equation}
  K = \frac{2 \pi t}{T}
  \label{eq:traj_k}
\end{equation}

In addition to the step height, the $x$ and $y$ translations, and the period, the robot's trajectory planner also considers two parameters to make the landing of the foot smoother. Since the robot’s actuators are servo motors controlled by position, the torque is proportional to the displacement it must perform between the points of the trajectory. Therefore, the higher the resolution of the trajectory, the smoother the movement. However, the resolution of the trajectory $N$ is fixed, given as a function of the step period and the frequency of the gait planner ($\SI{50}{\hertz}$) $N = 50T$. Thus, the strategy adopted is to distribute the $N$ points unevenly throughout the step period, so that there are more points near the moment the foot lands on the ground and less points near the moment the foot takes off. The $P_T$ parameter is a fraction of the total step period, and the $P_N$ is the fraction of the total number of points that should be between $0$ and $P_T \cdot T$. In other words, if $P_T = 0.66$ and $P_N = 0.33$, $33\%$ of $N$ will be in the first two-thirds of the period, while the remaining $67\%$ will be in the final one-third. The trajectory, considering these parameters, is illustrated in Fig. \ref{fig:traj_time_modified}.

@@

>>>>

INTRO

\subsection{Prototype}

A small quadruped was designed and developed for research (Fig. X).
During the prototyping phase, the hardware architecture and physics structure was designed. Following that, the structure of the robot was printed and integrated with the hardware components. The robot's actuators are servo motors of the model dynamixel MX-28, and a Raspberry Pi 4 was chosen as the processing unit. Furthermore, it contains an inertial sensor, model MPU6050. All the software was developed with the Robot Operating System 2 Humble (ROS 2) \cite{ROS2Humble}.

@@

The robot's structure is mammal-type, which is characterized by the presence of two joints in the sagittal plane, similar to animals such as dogs and horses.  The robot's legs follow a configuration called full-elbow, in which the four legs are oriented backward, as shown in Fig. X. Furthermore, it features 3 degrees of freedom (DoF) per leg, providing the legs with substantial freedom of movement. The robot was designed to favor the mass balance between the body and the robot's legs, that is, the majority of its mass is located in the body or as close to it as possible. Lighter legs can move quickly without significantly changing the center of gravity of the robot, which increases stability and requires less control complexity.  On the other hand, the legs should be strong enough to support the robot's weight and the impact on the ground.

The internal components, which include sensors, the processing unit, and the communication interface with the actuators were arranged symmetrically in the robot's body. The actuators (components that contribute to most of the system's mass) were positioned as close as possible to the body. The motor that acts on the tibia joint was placed in the upper part of the femur, to reduce the moment of inertia of the leg. This choice required the addition of a transmission system between the engine and the tibia link, consisting of a rigid metal rod with a ball joint at each edge.

\subsection{Gait Planner}

Quadruped robots move according to a gait pattern. The gait implemented for the robot follows the sequence of phases demonstrated in Fig. \ref{fig:trot_pattern}, in which the white rectangles represent the swing phase and the gray rectangles represent the stance phase. During the stance phase, the legs are on the ground and push the robot in the desired direction.  In the swing phase, the legs are raised and moved to the next support point. This gait pattern is called trot and is based on the concept of periodic and symmetrical gaits. The periodic gaits are characterized by the continuous repetition of the movements in the same instants. Symmetry is a characteristic of gaits that move a pair of legs together, leaving and returning to the ground in a synchronized way. Although the structure of the robot makes it possible to perform other types of gaits, in this work, only the trot was considered. To reduce the control complexity, a discontinuous gait was adopted, that is, the robot's body moves only when all the legs are on the ground, resulting in intermittent movement. In Fig. \ref{fig:trot_pattern}, it is noticeable that the same pair of diagonal legs always move at the same instant. Between two consecutive swing phases, there is a moment when all legs are in the stance phase, which is when the robot body is displaced in the desired direction of locomotion.

@@

<REESTRUTURAR ESSE PARAGRAFO>

The gait planner is responsible for controlling each leg and the robot's body. It defines where each foot will step and where the body should move based on the desired velocity command. Through a trajectory planner, it calculates the trajectory each foot and body should follow based on the velocity command. The gait planner then reads the pitch and roll angles of the body from the stabilization controller

The gait planner then sends commands to each joint based on the planned trajectories and the control efforts from the stabilization controller of the body at a frequency of $\SI{50}{\hertz}$. To convert both trajectories' points and the control efforts into the angles of the joints, a kinematic model is used, which is responsible for mapping the three-dimensional position of one foot to the joints' angles of its respective leg.

\subsection{Trajectory Planner}

The trajectory planner is responsible for calculating the trajectory that each foot must perform. For both the stance and swing phases, a cycloidal curve is adopted. As shown in \cite{Shi2021}, a cycloidal curve can be defined in the three-dimensional space by (\ref{eq:traj_x}) to (\ref{eq:traj_k})
between the points $(x_o, y_o, z_o)$ and $(x_f, y_f, z_f)$ as a function of time $t$, step height $H$, and period $T$.
\begin{equation}
  x = (x_f - x_o) \frac{K - \sin{(K)}}{2 \pi} + x_o
  \label{eq:traj_x}
\end{equation}
\begin{equation}
  y = (y_f - y_o) \frac{K - \sin{(K)}}{2 \pi} + y_o
  \label{eq:traj_y}
\end{equation}
\begin{equation}
  z = H \frac{1 - \cos{(K)}}{2} + z_o
  \label{eq:traj_z}
\end{equation}
\begin{equation}
  K = \frac{2 \pi t}{T}
  \label{eq:traj_k}
\end{equation}

In addition to the step height, the $x$ and $y$ translations, and the period, the robot's trajectory planner also considers two parameters to make the landing of the foot smoother. Since the robot’s actuators are servo motors controlled by position, the torque is proportional to the displacement it must perform between the points of the trajectory. Therefore, the higher the resolution of the trajectory, the smoother the movement. However, the resolution of the trajectory $N$ is fixed, given as a function of the step period and the frequency of the gait planner ($\SI{50}{\hertz}$) $N = 50T$. Thus, the strategy adopted is to distribute the $N$ points unevenly throughout the step period, so that there are more points near the moment the foot lands on the ground and fewer points near the moment the foot takes off. The $P_T$ parameter is a fraction of the total step period, and the $P_N$ is the fraction of the total number of points that should be between $0$ and $P_T \cdot T$. In other words, if $P_T = 0.66$ and $P_N = 0.33$, $33\%$ of $N$ will be in the first two-thirds of the period, while the remaining $67\%$ will be in the final one-third. The trajectory, considering these parameters, is illustrated in Fig. \ref{fig:traj_time_modified}.

\subsection{Kinematic Model}
\label{sec:detail_inv_kinematics}

@@

\subsection{Control Subsystem}

The control subsystem of the robot is composed of two main components: the joints' controllers, and the body's stabilization controller. The joints' controllers, each one a PID, are responsible for controlling each joint of the robot independently. Meanwhile, the body's stabilization controller is responsible for controlling the pitch and roll angles of the robot, maintaining it at 0 degrees.

<APROFUNDAR JOINT CONTROLLER>

The stabilization controller is formed by two PID controllers for the roll and pitch rotation of the robot's body. They act independently, stabilizing the rotation in each axis simultaneously. Both controllers are identical and were implemented following the block diagram in Fig. \ref{fig:pid}.

@@

The IMU sensor is responsible for measuring the rotation of the robot's body, providing feedback on the system's outputs. The limiter was added to avoid body rotations that exceed the limits of the joints.

@@

<COLOCAR OUTRO EXPERIMENTO>
